A full approximation scheme multilevel method \\ for nonlinear variational inequalities

### 2023-08-08 paper/fascd.tex
thank Max
I think this addresses the PF.

\title{A full approximation scheme multilevel method \\ for nonlinear variational inequalities\thanks{Draft \today.%Submitted to the editors DATE.
\funding{EB was supported by a Faculty Development Travel Award from United Academics, University of Alaska Fairbanks. PEF was supported by the Engineering and Physical Sciences Research Council [EPSRC grants EP/R029423/1 and EP/W026163/1]. This work used the ARCHER2 UK National Supercomputing Service (https://www.archer2.ac.uk). PEF thanks Lawrence Mitchell for advice on implementation, and Jack D.~Betteridge for assistance with Example \ref{ex:results:sia}.}}}

@@

Formula \eqref{eq:fe:levelsource} for $\ell^j$ is a classical FAS construction; compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.  It has the following explanation, which also justifies VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.  First, problem \eqref{eq:fe:downvi} was not solved exactly by the finer-level correction $y^{j+1}$---the down-smoother was only approximate---so we seek a further correction $y$ on the $j+1$ level such that
\begin{equation}
\ip{f^{j+1}(w^{j+1}+y^{j+1}+y)}{v-y^{j+1}-y} \ge \ip{\ell^{j+1}}{v-y^{j+1}-y} \quad \text{\emph{[notional]}} \label{eq:fe:downvinotional}
\end{equation}
for all $v\in \mathcal{D}^{j+1}$.  However, because the down-smoother has made progress, $y$ can be approximated by solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps: \emph{(i)} subtract the computable quantity $f^{j+1}(w^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides, \emph{(ii)} replace the residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction, \emph{(iii)} replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its injection, and \emph{(iv)} where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.  These steps yield a problem for the $j$th-level correction $y^j \in \mathcal{D}^j$:
\begin{align}
&\ip{f^j\left(\iR(w^{j+1}+y^{j+1})+y^j\right)}{v-y^j} - \ip{f^j\left(\iR(w^{j+1}+y^{j+1})\right)}{v-y^j} \label{eq:fe:downvicluttered} \\
&\qquad \ge \ip{R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right)}{v-y^j}, \notag
\end{align}
\pef{I don't see where in steps \emph{(i)-(iv)} the expression in the right of the angle brackets changes from $v - y^{j+1} - y$ (as in \eqref{eq:fe:downvinotional}) to $v - y^j$ (as in \eqref{eq:fe:downvicluttered}). From comparing the left hand sides, clearly $y^j$ in \eqref{eq:fe:downvicluttered} approximates $y$ in \eqref{eq:fe:downvinotional}, but from comparing the right hand sides, you want $y^j = y^{j+1} + y$. Maybe add a sentence, or another step, explaining what happened to the test expression?}
for all $v\in \mathcal{D}^{j}$.  This is VI problem \eqref{eq:fe:downvi}, but in cluttered form, motivating definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource}.  The explanation for problem \eqref{eq:fe:upvi} is similar.

@@

Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle with \id{down}, \id{up} smoother iterations on each level.  The $V(1,0)$ cycle, \id{up} $=0$, is similar to Algorithm 4.2 in \cite{GraeserKornhuber2009}, but generalized by applying FAS-type corrections and allowing any smoother or nonlinear operator $f$.  In unilateral cases one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} (Section \ref{sec:cd}) over CD \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over CDs \eqref{eq:fe:upwardsuminclusion}, and in this sense $V(1,1)=V(1,0)+V(0,1)$.

>>>>

\title{A full approximation scheme multilevel method \\ for nonlinear variational inequalities\thanks{Draft \today.%Submitted to the editors DATE.
\funding{EB was supported by a Faculty Development Travel Award from United Academics, University of Alaska Fairbanks, and thanks Max Heldman for helpful comments on the draft. PEF was supported by the Engineering and Physical Sciences Research Council [EPSRC grants EP/R029423/1 and EP/W026163/1]. This work used the ARCHER2 UK National Supercomputing Service (https://www.archer2.ac.uk). PEF thanks Lawrence Mitchell for advice on implementation, and Jack D.~Betteridge for assistance with Example \ref{ex:results:sia}.}}}

@@

Formula \eqref{eq:fe:levelsource} for $\ell^j$ is a classical FAS construction; compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.  It has the following explanation, inductively on VI problems \eqref{eq:fe:downvi}.  Because the down-smoother was only approximate, problem \eqref{eq:fe:downvi} was not solved exactly by the finer-level correction $y^{j+1} \in \mathcal{D}^{j+1}$.  Suppose we seek an improved correction by adding $y \in \mathcal{D}^{j}$, to give $y^{j+1}+y \in \mathcal{D}^{j+1}+\mathcal{D}^j \subset \cV^{j+1}$, so that
\begin{equation}
\ip{f^{j+1}(w^{j+1}+y^{j+1}+y)}{v-y} \ge \ip{\ell^{j+1}}{v-y} \quad \text{\emph{[notional]}} \label{eq:fe:downvinotional}
\end{equation}
for all $v \in \mathcal{D}^j$.  (Note $v-y = (y^{j+1}+v)-(y^{j+1}+y)$ and that nesting $\cV^j \subset \cV^{j+1}$ is assumed here.)

However, because the down-smoother has made progress, $y$ can be approximated by solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps: \emph{(i)} subtract the computable quantity $f^{j+1}(w^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides, \emph{(ii)} replace the relaxed residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction, \emph{(iii)} replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its injection, and \emph{(iv)} where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.  These steps yield a problem for the $j$th-level correction $y=y^j \in \mathcal{D}^j$:
\begin{align}
&\ip{f^j\left(\iR(w^{j+1}+y^{j+1})+y^j\right)}{v-y^j} - \ip{f^j\left(\iR(w^{j+1}+y^{j+1})\right)}{v-y^j} \label{eq:fe:downvicluttered} \\
&\qquad \ge \ip{R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right)}{v-y^j}, \notag
\end{align}
for all $v\in \mathcal{D}^{j}$.  Though in cluttered form, this is VI problem \eqref{eq:fe:downvi}, once definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource} are applied.

The explanation for \eqref{eq:fe:upvi} and \eqref{eq:fe:upwardinitial} is simpler.  Whether or not problem \eqref{eq:fe:downvi} has been exactly solved by the computed $y^j\in\mathcal{D}^j$, when going upward we can expand problem \eqref{eq:fe:downvi} into the larger set $\mathcal{U}^j \supset \mathcal{D}^j$.  (One may continue admissibly upward from any approximate solution to \eqref{eq:fe:upvi}.)  Generally $y^j$ is not the solution to \eqref{eq:fe:upvi}; in the optimization case $f=F'$ it would not be optimal in the enlarged constraint set.  From \eqref{eq:fe:upwardsuminclusion} our current best estimate of the solution to \eqref{eq:fe:upvi} is $z^j \approx y^j + Pz^{j-1}\in \mathcal{U}^j$; this gives the initial iterate for the up-smoother relaxation.

@@

Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle with \id{down} and \id{up} smoother iterations on a level, respectively.  The $V(1,0)$ cycle with \id{up} $=0$ is similar to Algorithm 4.2 in \cite{GraeserKornhuber2009}, but generalized by applying FAS-type corrections and allowing any smoother or nonlinear operator $f$.  In unilateral cases one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} (Section \ref{sec:cd}) over CD \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over CDs \eqref{eq:fe:upwardsuminclusion}, and in this sense $V(1,1)=V(1,0)+V(0,1)$.
